---
title: "Statistical Computing Workshop"
subtitle: "Northwestern University -- Political Science"
format: 
  html:
    toc: true
    title-block-banner: true
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(tinytable)
library(markdown)
```

## Spring 2025 Schedule

<!-- - {{< fa clock >}} &nbsp; 3:00 -- 4:20 PM -->
<!-- - {{< fa location-dot >}} &nbsp; Scott Hall 212 -->
- {{< fa video >}} &nbsp; [Zoom meeting](https://northwestern.zoom.us/j/94652528071?pwd=bfBocxKncYkfZbTYiCIlJqhqxwx7EY.1)
- {{< fa file-pdf >}} &nbsp; [Full AY 2024-25 schedule](schedule.pdf)
- {{< fa file-excel >}} &nbsp; [Attendance sheet for registered students](https://forms.office.com/r/m8ZT8aDpVr)

```{r schedule, include = FALSE}
schedule =
  tribble(
    ~Date, ~Time, ~Topic, ~Presenter,
    "Tuesday, April 29", "9:30 -- 10:50 AM", "[Spatial Data](spatial/)", "[Nicole Wilson](https://www.nicoleewilson.com)",
    "Tuesday, May 13", "9:30 -- 10:50 AM", "[Images as Data](images/)", "[Hyein Ko](https://www.heyikon.com)",
    "Tuesday, June 3", "9:30 -- 10:50 AM", "Coding session", " "
    )
```


```{r show-schedule, echo = FALSE, results = "asis"}
schedule %>% 
  tt(width = 1) %>% 
  format_tt(j = 2:4, markdown = TRUE)
```


## Details

The Statistical Computing Workshop (formerly R Workshop) is a year-long series that meets three times per quarter during the academic year. The purpose of the workshop is to learn, practice, and update cutting edge statistical programming skills as they apply to quantitative and computational social science. 

Workshop meetings will feature internal or external speakers introducing a new tool, method, approach, or practice. All meetings are hybrid or fully virtual.

The workshop is open to everyone, but those formally enrolled will receive full credits by attending to every meeting. To receive credit, you must sign up to the corresponding version of POLI_SCI 490 in the **Spring** quarter.

Contact [Gustavo Diaz](https://gustavodiaz.org/) for additional information.

## Previous Workshops

```{r previous, include = FALSE}
previous = tribble(
  ~Topic, ~Presenter, ~`Last offered`,
  "[Geographical and Census Data](census/)", "[Michelle Bueno Vasquez](https://michellebuenovasquez.wordpress.com)", "Winter 2025",
  "[Interpreting Statistical Models](models/)", "[Vincent Arel-Bundock](https://arelbundock.com)", "Winter 2025",
  "[Presentations and Posters](presentations/)", "[Gustavo Diaz](https://gustavodiaz.org/)", "Fall 2024",
  "[Writing papers in Quarto](papers/)", "[Gustavo Diaz](https://gustavodiaz.org/)","Fall 2024"
)
```

```{r show-previous, echo = FALSE, results = "asis"}
previous %>% 
  tt(width = 1) %>% 
  format_tt(j = 1:2, markdown = TRUE)
```



