---
title: "Statistical Computing Workshop"
subtitle: "Northwestern University -- Political Science"
format: 
  html:
    toc: true
    title-block-banner: true
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(tinytable)
library(markdown)
```

## Spring 2026 Schedule

- {{< fa clock >}} &nbsp; 12:30 -- 2:00 PM
- {{< fa location-dot >}} &nbsp; Hybrid (see event links for location)
- {{< fa file-pdf >}} &nbsp; [Full AY 2025-26 schedule](schedule.pdf)
- {{< fa file-excel >}} &nbsp; [Attendance sheet for registered students](https://forms.office.com/r/m8ZT8aDpVr)

```{r schedule, include = FALSE}
schedule =
  tribble(
    ~Date, ~Topic, ~Presenter,
    "Wednesday, April 15", "[Estimating heterogeneous causal effects with varying coefficient models and Bayesian Additive Regression Trees](https://planitpurple.northwestern.edu/event/640955)", "Sameer Deshpande",
    "Wednesday, April 22", "[Leveraging Python and JAX in R workflows](https://planitpurple.northwestern.edu/event/640958)", "Andrés Cruz",
    "TBD", "TBD", "TBD"
    )
```


```{r show-schedule, echo = FALSE, results = "asis"}
schedule %>% 
  tt(width = 1) %>% 
  format_tt(j = 2:3, markdown = TRUE)
```


## Details

The Statistical Computing Workshop is a year-long series that meets three times per quarter during the academic year. The purpose of the workshop is to learn, practice, and update cutting edge statistical programming skills as they apply to quantitative and computational social science. 

Workshop meetings will feature internal or external speakers introducing a new tool, method, or research project involving statistical computing in the broadest sense. All meetings are hybrid or fully virtual.

The workshop is open to everyone. Graduate students in political science formally enrolled in the corresponding version of POLI_SCI 490 will receive full credit by attending every meeting. 

Contact [Gustavo Diaz](https://gustavodiaz.org/) for additional information.

## Previous Workshops

```{r previous, include = FALSE}
previous = tribble(
  ~Topic, ~Presenter, ~`Term`,
  "[Hands-On Methods for Text-as-Data Research](https://planitpurple.northwestern.edu/event/636981)", "[Artur Baranov](https://artur-baranov.github.io)", "Winter 2026",
  "[Building reproducible workflows with `{targets}`](https://planitpurple.northwestern.edu/event/636983)", "[Andrew Heiss](https://www.andrewheiss.com)", "Winter 2026",
  "[Rethinking Computational Multimodal Analysis in Social Science Research](https://planitpurple.northwestern.edu/event/636982)", "[Yingdan Lu](https://yingdanlu.com)", "Winter 2026",
  "[Dos and Don'ts of Coding with AI](coding_ai/)", "[Gustavo Diaz](https://gustavodiaz.org)", "Fall 2025",
  "[A Practical Guide to Ranking Data Analysis in the Social Sciences](ranking/)", "[Yuki Atsusaka](https://atsusaka.org)", "Fall 2025",
  "Declaring and Diagnosing Research Designs", "[Alexander Coppock](https://alexandercoppock.com)", "Fall 2025",
  "[Images as Data](images/)", "[Hyein Ko](https://www.heyikon.com/)",
  "Spring 2025",
  "[Spatial Data](spatial/)", 
  "[Nicole Wilson](https://www.nicoleewilson.com/)", "Spring 2025",
  "[Geographical and Census Data](census/)", "[Michelle Bueno Vasquez](https://michellebuenovasquez.wordpress.com)", "Winter 2025",
  "[Interpreting Statistical Models](models/)", "[Vincent Arel-Bundock](https://arelbundock.com)", "Winter 2025",
  "[Presentations and Posters](presentations/)", "[Gustavo Diaz](https://gustavodiaz.org/)", "Fall 2024",
  "[Writing papers in Quarto](papers/)", "[Gustavo Diaz](https://gustavodiaz.org/)","Fall 2024"
)
```

```{r show-previous, echo = FALSE, results = "asis"}
previous %>% 
  tt(width = 1) %>% 
  format_tt(j = 1:2, markdown = TRUE)
```



